Statistical Reporting Inconsistencies in Experimental Linguistics

### 2025-05-20 manuscript/manuscript_anonymized.tex
ready for resubmission

\author{anonymous}

@@

\leftheader{anonymous}

\abstract{The present paper investigates the prevalence of statistical
reporting consistencies across articles in eight experimental
linguistics journals published between 2000 and 2023. Using the R
package ``Statcheck'', we retrieved 85,442 statistical test from 13,065
articles and assessed whether p-values were consistent with their test
statistic and degrees of freedom. Half of all articles (50\%) that used
null hypothesis significance testing contained at least one inconsistent
p-value. More than one in eight articles (13\%) contained an
inconsistency that may have affected the statistical conclusion. The
inconsistency rates were comparable across journals and seem stable over
publication years. We discuss possible reasons for this high rate and
offer actionable steps for authors, reviewers, and editors to remedy
this state-of-affairs.}

\keywords{statistics, statcheck, reproducibility, null hypothesis
significance testing}

\authornote{ 
\par{ }
\par{   The authors have no conflict of interest to declare.    }
\par{Correspondence concerning this article should be addressed to }
}

@@

\section{1. Introduction}\label{introduction}

What we know about human language and its cognitive underpinnings is
often informed by experimental data. Researchers test theoretical
predictions with their data using statistical tests. Depending on the
results of their tests, researchers make claims for or against
theoretical assumptions.\\
Since these tests play such an integral part in the argumentation
process of experimentalists, both the data these tests are based on and
the computational procedure of the tests itself should be transparent in
order for other researchers to critically evaluate them. But more than
that, things can go wrong. Humans are fallible; they make mistakes.
Transparent sharing allows others to detect and correct human error. In
recent years, quantitative linguistics have seen repeated calls to
become more transparent and reproducible by sharing data and statistical
protocols, often under the banner of ``open science''
(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
calls, sharing of statistical protocols is still rather rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}). If statistical procedures cannot be critically evaluated,
human errors might be left undetected and thus remain uncorrected in the
publication record. And if undetected errors affect the decision
procedure of the analysis, i.e.~whether a hypothesis is rejected or
accepted, these errors might lead to -at best- overconfident, -at worst-
false theoretical conclusions. The present paper will present evidence
that the published literature in experimental linguistics contains a
concerning amount of such statistical errors, a state of affairs which
warrants more rigorous data sharing practices.

\section{2. Statistal reporting
inconsistencies}\label{statistal-reporting-inconsistencies}

@@

To assess the prevalence of statistical-reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
p-values reported in thirteen experimental linguistic journals published
between 2000 and 2023. We explore whether the rate of inconsistency
differs across journals, whether their rate has changed over the course
of the last 23 years and whether there is evidence for bias in these
statistical-reporting inconsistencies. We discuss the results and offer
concrete recommendations for authors, reviewers, and editors to tackle
this problem.

\section{3. Method}\label{method}

\subsection{3.1. Data availability
statement}\label{data-availability-statement}

@@

\subsection{3.2. Sample}\label{sample}

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the search string ``experiment*'' in title, abstract
and/or keywords. Out of these 100 journals, we selected all journals
with at least 10\% of articles containing the search string
``experiment*''. Out of the remaining 37 journals, we selected those
journals that urged APA formatting either in the main body of the text
or specifically regarding statistics in the author guidelines, resulting
in nine remaining journals. Moreover, to access the articles in .pdf
format, the articles had to be either accessible to us through our
library license, or open access, resulting in a final list of eight
journals: Applied Psycholinguistics (APS), Bilingualism: Language and
Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
and Speech (LaS), Language Learning and Techology (LLT), Journal of
Language and Social Psychology (LSP), Journal of Child Language (JCL),
and Studies in Second Language Acquisition (SLA). An anonymous reviewer
raised the justified concern that the resulting sample might not
represent the vast majority of work in experimental linguistics. To
address this concern, we additionally included those five journals that
have published the highest absolute number of experimental articles,
regardless of whether they explicitly urge APA formatting or not,
resulting in the inclusion of the following psycholinguistic outlets:
Journal of Memory and Language (JML), Language Cognition and
Neuroscience (LCN, former Language and Cognitive Processes), Journal of
Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing
Research (SLH), and Brain and Language (BAL).

@@

\subsection{3.3. Statcheck}\label{statcheck}

We used the R package Statcheck (Version 1.5.0,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
automatically detect statistical-reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using regular expressions. That way, Statcheck can detect
results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and Q
tests as long as the test result fulfills three conditions: (a) the test
result is reported completely including the test statistic, the degrees
of freedom (if applicable), and the p-value; (b) the test result is in
the body of the text, i.e. Statcheck usually misses information in
tables; and (c) the test result is reported in American Psychological
Association style (\citeproc{ref-APA2020}{American Psychological
Association, 2020}). Given these constraints, Statcheck is estimated to
detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-value and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

Articles from LAB spanned 2011-2023; LCN spanned 2015-2023. Statcheck
could not parse 291 articles, likely related to issues with rendering
the Chi-Squared symbol being erroneously converted from .pdf to .txt,
resulting in some gaps in the coverage. This procedure resulted in 13065
research articles which were submitted to analysis.

\subsection{3.4. Analysis and research
questions}\label{analysis-and-research-questions}

The aims of this study were explicitly exploratory and hypothesis
generating, thus analyses remain merely descriptive by reporting on the
proportion of articles/tests that are statistically (in)consistent. Our
investigation was set out to explore the following research questions:
(1) How prevalent are statistical inconsistencies (and decision
inconsistencies) in our sample? (2) Do inconsistency rates vary across
journals and/or their publication year? And (3) is there evidence for
bias, i.e.~are processes that result in inconsistencies more likely to
produce lower p-values?

\section{4. Results}\label{results}

\subsection{4.1 Inconsistencies are highly
prevalent}\label{inconsistencies-are-highly-prevalent}

@@

\subsection{4.2 Inconsistencies are stable across journals and
publication
year}\label{inconsistencies-are-stable-across-journals-and-publication-year}

@@

\subsection{4.3 Inconsistencies are
biased}\label{inconsistencies-are-biased}

Inconsistencies that report the p-value as being larger or smaller than
a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05,
respectively) are not equally prevalent in the sample: There were 4.4\%
of inconsistencies with p being reported as larger than a reference but
7.1\% inconsistencies with p being reported as smaller than a reference.
So even if we assumed these inconsistencies were merely typos of the
comparison sign (e.g.~\textless{} instead of \textgreater),
inconsistencies that erroneously report the p-value to be smaller than a
reference value are more frequent than inconsistencies that erroneously
report the p-value to be larger than a reference value.

These biases are also reflected in decision inconsistencies. Of all
decision inconsistencies (n = 1442), 72\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05), i.e.~non-significant results
are more than twice as likely to be erroneously reported as significant
than the other way around. The latter pattern, however, seems to be a
bias of the past. Reproducing Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
plots the development of the bias observed for decision inconsistencies
over time. The prevalence of decision inconsistencies in significant
p-values seems to have slightly decreased over the years, while the
prevalence of decision inconsistencies in non-significant p-values seems
to have slightly increased over the years.

@@

\section{5. Discussion and
Recommendations}\label{discussion-and-recommendations}

\subsection{5.1. Statistical reporting inconsistencies are
prevalent}\label{statistical-reporting-inconsistencies-are-prevalent}

The present study found a large amount of statistical reporting
inconsistencies across a sample of 13065 experimental linguistic
articles, containing 85442 assessable p-values. 13.2\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency.

The present study can be considered a conceptual replication of previous
studies investigating statistical-reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent assessments using Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
discovered inconsistency rates fall in line with these studies that
report on rates between 4\% and 14\%, with between 10\% and 63\% of
articles containing at least one inconsistency.

@@

The present examination did not indicate any noticeable trend over a
period of 23 years of publications. Observed inconsistencies were
characterized by reported p-values being on average lower than their
recalculated counterparts and the prevalence of decision inconsistencies
was higher for p-values reported as significant than for those reported
as non-significant. These biases have been reported on for other
disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}).
These skewed patterns could indicate a systematic bias in favor of lower
p-values in general and a bias towards significant results in
particular. Our data do not speak to the causes of these biases, but
possible reasons include the following:

@@

Third, the observed bias might merely be a reflection of publication
bias (\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}) with
(erroneously) reported significant p-values being more likely to be
published than non-significant ones. Publication bias is a well
established pattern in experimental linguistic research with many recent
meta analyses discussing possible evidence for it
(\citeproc{ref-isbilen2022statistical}{Isbilen \& Christiansen, 2022};
\citeproc{ref-lehtonen2018bilingualism}{Lehtonen et al., 2018};
\citeproc{ref-lu2024meta}{Lu et al., 2024}) For example, De Bruin et al.
(\citeproc{ref-de2015cognitive}{2015}) showed that studies with results
supporting the bilingual-advantage theory were most likely to be
published, while studies with results challenging the theory were
significantly less likely to be published.

Regardless of possible causes for bias in the processes that generate
inconsistencies, the data also suggest that some of these biases has
decreased over time, pointing to a positive development.

\subsection{5.2. Limitations of our
study}\label{limitations-of-our-study}

@@

\subsection{5.3. Recommendations for the
field}\label{recommendations-for-the-field}

@@

Journal editors could explicitly recommend consistency checks with
algorithms such as Statcheck during peer review, a practice that has
been taken up on by several journals from neighboring disciplines
(Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
  accessed on July 15, 2024.}, Advances in Methods and Practices in
Psychological Science\footnote{https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines;
  accessed on on July 15, 2024.}, Stress \& Health Barber
(\citeproc{ref-barber2017meticulous}{2017})). Editors could also demand,
recommend or at least encourage data sharing for publication in their
journal. Data sharing policies have been shown to substantially increase
the reproducibility of analyses (e.g.,
\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022})
and a number of linguistic journals have already implemented such
policies, including journals within our sample. Having said that, open
data policies which for example the Journal of Memory and Language
introduced in 2018 did not seem to have affected the proportion of
statistical inconsistencies. The inconsistency rates for JML (and other
journals) were rather stable across time, so open data alone might not
resolve the issue without further changes to the research eco-system.

>>>>

\authorsnames{anonymous,anonymous}

@@

\leftheader{anonymous and anonymous}

\abstract{The present paper investigates the prevalence of statistical
reporting inconsistencies across articles in thirteen experimental
linguistics journals published between 2000 and 2023. Using the R
package ``Statcheck'', we retrieved 85,442 statistical test from 13,065
articles and assessed whether p-values were consistent with their test
statistic and degrees of freedom. Half of all articles (50\%) that used
null hypothesis significance testing contained at least one inconsistent
p-value. More than one in eight articles (13\%) contained an
inconsistency that may have affected the statistical conclusion. The
inconsistency rates were comparable across journals and seem stable over
publication years. We discuss possible reasons for this high rate and
offer actionable steps for authors, reviewers, and editors to remedy
this state-of-affairs.}

\keywords{statistics, Statcheck, reproducibility, null hypothesis
significance testing}

\authornote{

\par{       }
\par{Correspondence concerning this article should be addressed to }
}

@@

\section{Introduction}\label{introduction}

What we know about human language and its cognitive underpinnings is
often informed by experimental data. Researchers test theoretical
predictions with their data using statistical tests. Depending on the
results of their tests, researchers make claims for or against
theoretical assumptions.\\
Since these tests play such an integral part in the argumentation
process of experimentalists, both the data these tests are based on and
the computational procedure of the tests itself should be error-free.
But humans are fallible; they make mistakes. We cannot avoid making
errors, but we can make them at least detectable. Transparent sharing
allows others to detect and correct human error.

In recent years, quantitative linguistics have seen repeated calls to
become more transparent and reproducible by sharing data and statistical
protocols, often under the banner of ``open science''
(\citeproc{ref-arvan2022reproducibility}{Arvan et al., 2022};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022};
\citeproc{ref-roettger2019researcher}{Roettger, 2019}). Despite these
calls, sharing of statistical protocols is still rather rare across the
language sciences (\citeproc{ref-bochynska2023reproducible}{Bochynska et
al., 2023}). If statistical procedures cannot be critically evaluated,
human errors might be left undetected and thus remain uncorrected in the
publication record. And if undetected errors affect the decision
procedure of the analysis, i.e.~whether a hypothesis is rejected or
accepted, these errors might lead to -at best- overconfident, -at worst-
false theoretical conclusions. The present paper will present evidence
that the published literature in experimental linguistics contains a
concerning amount of such statistical inconsitencies, a state of affairs
which warrants more rigorous data sharing practices.

\section{Statistal reporting
inconsistencies}\label{statistal-reporting-inconsistencies}

@@

To assess the prevalence of statistical reporting inconsistencies in
experimental linguistics, the present paper conceptually replicates
Nuijten et al. (\citeproc{ref-nuijten2016prevalence}{2016}) and assesses
p-values reported in thirteen experimental linguistic journals published
between 2000 and 2023. We explore whether the inconsistency rates differ
across journals, whether they have changed over the course of the last
23 years and whether there is evidence for bias in these
inconsistencies. We discuss the results and offer concrete
recommendations for authors, reviewers, and editors to tackle this
problem.

\section{Method}\label{method}

\subsection{Data availability
statement}\label{data-availability-statement}

@@

\subsection{Sample}\label{sample}

Focusing on experimental linguistic research, we used Kobrock and
Roettger (\citeproc{ref-kobrock2023assessing}{2023}) as a point of
departure who list 100 linguistic journals that had at least a hundred
articles published at the time of assessment (2021) and a high ratio of
articles containing the search string ``experiment*'' in title, abstract
and/or keywords. Out of these 100 journals, we selected all journals
with at least 10\% of articles containing the search string
``experiment*''. Out of the remaining 37 journals, we selected those
journals that urged APA formatting either in the main body of the text
or specifically regarding statistics in the author guidelines, resulting
in nine remaining journals. Moreover, to access the articles in .pdf
format, the articles had to be either accessible to us through our
library license, or open access, resulting in a final list of eight
journals: Applied Psycholinguistics (APS), Bilingualism: Language and
Cognition (BLC), Linguistic Approaches to Bilingualism (LAB), Language
and Speech (LaS), Language Learning and Techology (LLT), Journal of
Language and Social Psychology (LSP), Journal of Child Language (JCL),
and Studies in Second Language Acquisition (SLA). An anonymous reviewer
raised the justified concern that the resulting sample might not
represent the vast majority of work in experimental linguistics. To
address this concern, we additionally included those five journals that
have published the highest absolute number of experimental articles
(according to \citeproc{ref-kobrock2023assessing}{Kobrock \& Roettger,
2023}), regardless of whether they explicitly urge APA formatting or
not, resulting in the inclusion of the following psycholinguistic
outlets: Journal of Memory and Language (JML), Language, Cognition and
Neuroscience (LCN, former Language and Cognitive Processes), Journal of
Psycholinguistic Research (JPR), Journal Of Speech Language And Hearing
Research (SLH), and Brain and Language (BAL).

@@

\subsection{Statcheck}\label{statcheck}

We used the R package Statcheck (Version 1.5.0,
\citeproc{ref-nuijten2023statcheck}{Nuijten \& Epskamp, 2024}) to
automatically detect statistical reporting inconsistencies. Statcheck
works as follows: After converting articles in pdf or html format to
plain text, Statcheck searches for specific strings that correspond to a
NHST result using regular expressions. That way, Statcheck can detect
results of t-tests, F-tests, Z-tests, χ2-tests, correlation tests, and
Q-tests as long as the test result fulfills three conditions: (a) the
test result is reported completely including the test statistic, the
degrees of freedom (if applicable), and the p-value; (b) the test result
is in the body of the text, i.e. Statcheck usually misses information in
tables; and (c) the test result is reported in American Psychological
Association style (\citeproc{ref-APA2020}{American Psychological
Association, 2020}). Given these constraints, Statcheck is estimated to
detect roughly 60\% of all reported NHST results
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). Statcheck
uses the reported test statistic and degrees of freedom to recalculate
the p-value, compares the reported and recalculated p-value and, if
there is a mismatch, flags the test as containing an ``inconsistency.''
The algorithm takes into account that tests might have been performed as
one-tailed by identifying the search strings ``one-tailed,''
``one-sided,'' or ``directional'' in the body of the text. Moreover,
Statcheck considers p = .000 and p \textless{} .000 as inconsistent
because p-values of exactly zero are mathematically impossible and the
APA manual (\citeproc{ref-APA2020}{American Psychological Association,
2020}) advises to report very small p-values as p \textless{} .001.
Validity checks of Statcheck suggest that inter-rater reliability
between manual coding and Statcheck is high, i.e.~0.76 for
inconsistencies and 0.89 for decision inconsistencies
(\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
overall accuracy of Statcheck is estimated to be between 96.2\% to
99.9\% (\citeproc{ref-nuijten2017validity}{Nuijten et al., 2017}). We
thus consider Statcheck a valid proxy of the prevalence of statistical
reporting inconsistencies.

Articles from Linguistic Approaches to Bilingualism spanned 2011-2023;
articles from Language, Cognition and Neuroscience spanned 2015-2023.
Statcheck could not parse 291 articles, likely related to issues with
rendering the Chi-Squared symbol being erroneously converted from .pdf
to .txt, resulting in some gaps in the coverage. This procedure resulted
in 13065 research articles which were submitted to analysis.

\subsection{Analysis and research
questions}\label{analysis-and-research-questions}

The aims of this study were explicitly exploratory and hypothesis
generating, thus analyses remain merely descriptive by reporting on the
proportion of articles/tests that are statistically (in)consistent. Our
investigation was set out to explore the following research questions:
(1) How prevalent are statistical inconsistencies (and decision
inconsistencies) in our sample? (see Section~\ref{sec-prevalent}) (2) Do
inconsistency rates vary across journals and/or their publication year?
(see Section~\ref{sec-stable}) And (3) is there evidence for bias,
i.e.~are processes that result in inconsistencies more likely to produce
lower p-values? (see Section~\ref{sec-bias})

\section{Results}\label{results}

\subsection{Inconsistencies are highly prevalent}\label{sec-prevalent}

@@

\subsection{Inconsistencies are stable across journals and publication
year}\label{sec-stable}

@@

\subsection{Inconsistencies appear biased but bias decreases over
time}\label{sec-bias}

If inconsistencies were bias-free, we would expect different types of
inconsistencies to be equally frequent. However, this is not the case.
Inconsistencies that report the p-value as being larger or smaller than
a reference value (e.g.~p \textgreater{} 0.05 or p \textless{} 0.05,
respectively) are not equally prevalent in the sample: There were 4.4\%
of inconsistencies with p being reported as larger than a reference but
7.1\% inconsistencies with p being reported as smaller than a reference.
So even if we assumed these inconsistencies were merely typos of the
comparison sign (e.g.~\textless{} instead of \textgreater),
inconsistencies that erroneously report the p-value to be smaller than a
reference value are more frequent than inconsistencies that erroneously
report the p-value to be larger than a reference value.

These biases are also reflected in decision inconsistencies. Of all
decision inconsistencies (n = 1442), 72\% represent cases in which a
reported significant result (p \textless{} 0.05) is recalculated as
non-significant (p \textgreater{} 0.05), i.e.~non-significant results
are more than twice as likely to be erroneously reported as significant
than the other way around. The latter pattern, however, seems to have
decreased over time. Reproducing Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}), Figure~\ref{fig-scatter}
plots the development of the bias observed for decision inconsistencies
over time. The prevalence of decision inconsistencies in significant
p-values seems to have slightly decreased over the years, while the
prevalence of decision inconsistencies in non-significant p-values seems
to have slightly increased over the years.

@@

\section{Discussion and
Recommendations}\label{discussion-and-recommendations}

\subsection{Statistical reporting inconsistencies are
prevalent}\label{statistical-reporting-inconsistencies-are-prevalent}

The present study found a large amount of statistical reporting
inconsistencies across a sample of 13065 experimental linguistic
articles, containing 85442 assessable p-values. 13.2\% of all p-values
were flagged as inconsistent and 1.7\% were flagged as decision
inconsistencies, i.e.~the reported p-value is on the opposite side of
the alpha threshold than the recalculated p-value. On average, 50\% of
assessable articles contained at least one inconsistency and 13\%
contained at least one decision inconsistency. The present examination
did not indicate any noticeable trend for inconsistency rates across
journals or publication year.

The present study can be considered a conceptual replication of previous
studies investigating statistical reporting inconsistency across
different disciplines (\citeproc{ref-bakker2011mis}{Bakker \& Wicherts,
2011}, \citeproc{ref-bakker2014outlier}{2014};
\citeproc{ref-caperos2013consistency}{Caperos \& Pardo, 2013};
\citeproc{ref-veldkamp2014statistical}{Veldkamp et al., 2014};
\citeproc{ref-wicherts2011willingness}{Wicherts et al., 2011}) and most
recent assessments using the automatic tool Statcheck
(\citeproc{ref-buckley2023estimating}{Buckley et al., 2023};
\citeproc{ref-colombo2018statistical}{Colombo et al., 2018};
\citeproc{ref-gross2021fidelity}{Groß, 2021};
\citeproc{ref-nuijten2016prevalence}{Nuijten et al., 2016}). The
discovered inconsistency rates fall in line with these studies that
report on rates between 4\% and 14\%, with between 10\% and 63\% of
articles containing at least one inconsistency. Moreover, the observed
rates of inconsistencies and decision inconsistencies are virtually
identical to rates reported on by Nuijten et al.
(\citeproc{ref-nuijten2016prevalence}{2016}) on the psychological
science literature.

@@

Our results suggest biases as well. The prevalence of decision
inconsistencies was higher for p-values reported as significant than for
those reported as non-significant. These biases have been reported on
for other disciplines as well (\citeproc{ref-bakker2011mis}{Bakker \&
Wicherts, 2011}; \citeproc{ref-nuijten2016prevalence}{Nuijten et al.,
2016}). These skewed patterns could indicate a systematic bias in favor
of lower p-values in general and a bias towards significant results in
particular. Our data do not speak to the causes of these biases, but
possible reasons include the following:

@@

Third, the observed bias might merely be a reflection of publication
bias (\citeproc{ref-franco2014publication}{Franco et al., 2014};
\citeproc{ref-sterling1959publication}{Sterling, 1959}) with
(erroneously) reported significant p-values being more likely to be
published than (erroneously) reported non-significant ones. Publication
bias is a well established pattern in experimental linguistic research
with many recent meta analyses discussing possible evidence for it
(\citeproc{ref-isbilen2022statistical}{Isbilen \& Christiansen, 2022};
\citeproc{ref-lehtonen2018bilingualism}{Lehtonen et al., 2018};
\citeproc{ref-lu2024meta}{Lu et al., 2024}) For example, De Bruin et al.
(\citeproc{ref-de2015cognitive}{2015}) showed that studies with results
supporting the bilingual-advantage theory were most likely to be
published, while studies with results challenging the theory were
significantly less likely to be published.

Regardless of what possibly causes biases in the processes that generate
inconsistencies, our data also suggest a positive development. The over
proportional occurrence of decision inconsistencies for p-values
reported as significant has decreased over time.

\subsection{Limitations of our study}\label{limitations-of-our-study}

@@

\subsection{Recommendations for the
field}\label{recommendations-for-the-field}

@@

Journal editors could explicitly recommend consistency checks with
algorithms such as Statcheck during peer review, a practice that has
been taken up on by several journals from neighboring disciplines
(Psychological Science\footnote{http://www.psychologicalscience.org/publications/psychological\_science/ps-submissions;
  accessed on July 15, 2024.}, Advances in Methods and Practices in
Psychological Science\footnote{https://www.psychologicalscience.org/publications/ampps/ampps-submission-guidelines;
  accessed on on July 15, 2024.}, Stress \& Health Barber
(\citeproc{ref-barber2017meticulous}{2017})). Editors could also demand,
recommend or at least encourage data sharing for publication in their
journal. Data sharing policies have been shown to substantially increase
the reproducibility of analyses (e.g.,
\citeproc{ref-hardwicke2018data}{Hardwicke et al., 2018};
\citeproc{ref-laurinavichyute2022share}{Laurinavichyute et al., 2022})
and a number of linguistic journals have already implemented such
policies, including journals within our sample. Having said that, open
data policies which for example the Journal of Memory and Language
introduced in 2018 did not seem to have affected the proportion of
statistical inconsistencies after their introduction. The inconsistency
rates for JML (and other journals) were rather stable across time, so
open data alone might not resolve the issue without further changes to
the research eco-system.
